---
title: "Excercise 3: Logistic regression"
output: 
  html_document: 
    keep_md: true 
    code_folding: hide 
---

##  1. Load data
  
##  2. Correlations in data
  
```{r}
library(ISLR)
library(corrplot)

data("Smarket")
names(Smarket)
summary(Smarket)
cor(Smarket[,-9])
corrplot(cor(Smarket[,-9]), method = "number", type = "upper")

```

##  3. plot Volume
  
```{r}
par(mfrow = c(1,2))
plot(Smarket$Volume)
plot(Smarket$Volume~ as.factor(Smarket$Year), xlab = "Year", ylab = "Volume")

```
  
##  4. Fit glm() with logistic regression 
### (Y = Direction, X = Lag[1-5], Volume)
  
```{r}
glm.fit = glm(Direction ~ 
                Lag1 + Lag2 +Lag3 +Lag4 +Lag5 +Volume,
              data = Smarket, 
              family = binomial)
summary(glm.fit)
coef(glm.fit)
cat("following is p-values column")
summary(glm.fit)$coef[,4]

```
  
##  5. Predict probabilities for Direction (1 = up, 2 = down)
###  Type = "response" [P(Y = 1|X)]
  
##  6. Make predictions for Up and down separately
  
##  7. Create confusion matrix 
### (correctly or incorrectly classified directions in predictions)

```{r}
contrasts(Smarket$Direction)

glm.prob = predict(glm.fit, type = "response")
glm.prob[1:10]
glm.pred <- rep("Down", length(glm.prob))
glm.pred[glm.prob > .5] <- "Up"
knitr::kable(table(glm.pred, Smarket$Direction), caption = "confusion matrix")
cat(paste0("correct predictions = ", (507+145)/1250))
mean(glm.pred == Smarket$Direction)
cat(paste0("training error rate  = [100 - 52.2]% = ", 100-52.2, "%"))
```
  
##  8. Split data in training and test
  
```{r}
train = (Smarket$Year < 2005)
Smarket_2005 <- Smarket[!train,]
Direction_2005 <- Smarket$Direction[!train]
```
  
##  9. glm() fit using subset
  
##  10. Look at predictions on test
  
```{r}
glm.fit2 <- glm(Direction ~
                  Lag1 + Lag2 + Lag3 + Lag4 + Lag5 + Volume,
                data = Smarket, subset = train, 
                family = binomial)

glm.prob2 <- predict(glm.fit2, Smarket_2005, type = "response")
glm.pred2 <- rep("Down", length(glm.prob2))
glm.pred2[glm.prob2 > 0.5] <- "Up"
knitr::kable(table(glm.pred2, Direction_2005), caption = "prediction error")
cat(paste0("the prediction accuracy on test = ", mean(glm.pred2 == Direction_2005)))
cat(paste0("the test error rate = ", mean(glm.pred2 != Direction_2005)))
```
  
##  11. Refit glm() with only `Lag1` and `Lag2`

```{r}
glm.fit3=glm(Direction~ 
             Lag1+Lag2,
             data=Smarket, subset=train, 
             family=binomial)
summary(glm.fit3)
glm.prob3=predict(glm.fit3,Smarket_2005,type="response")
glm.pred3=rep("Down",252)
glm.pred3[glm.prob3>.5]="Up"
knitr::kable(table(glm.pred3,Direction_2005), caption = "prediction error")
cat(paste0("prediction accuracy on test = ", mean(glm.pred3==Direction_2005)))
cat(paste0("true positives for Up = ", 106/(106+76)))
```
  
##  12. Predict Direction 
### when `Lag1 = 1.2,1.5` and `Lag2 = 1.1,-0.8`
  
```{r}
data.frame(Lag1 = c(1.2,1.5), Lag2 = c(1.1,-0.8))
predict3 <- predict(glm.fit3,
        newdata = data.frame(Lag1 = c(1.2,1.5), Lag2 = c(1.1,-0.8)), 
        type = "response")
cbind(data.frame(Lag1 = c(1.2,1.5), Lag2 = c(1.1,-0.8)), pred = predict3)

```
  
  